---
title: "Appendices"
author: "Group 34"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include = FALSE}
# Load any packages you may need here
library(ggplot2)
library(gridExtra)
options(repos = c(CRAN = "https://cloud.r-project.org"))
install.packages("car")
library(car)
 # can use if you want to create side-by-side plots (like in Figure 5)
```

```{r readData, include = FALSE}
# Save this template file in your final project's RStudio Project 
# Edit this code chunk to read in your data

cancer_mortality_rates <- read.csv("cancer_mortality_rates.csv")
```

------------------------------------------------------------------------

#### PART 1 : Figure 1.1: Summary Ouput - Initial Model

```{r figure1.1, echo = FALSE}
# Be sure to use echo = FALSE in each code chunk to suppress your code! Your Appendices file should not include any code, just output!
lm_initial <- lm(cmRate ~ medianAge + pctPoverty + pctBach + rmRace, data = cancer_mortality_rates)
summary(lm_initial)
```

------------------------------------------------------------------------

#### Figure 1.2: Residuals Plot - Initial Model

```{r figure1.2, echo = FALSE}
plot(lm_initial, which = 1)
```

------------------------------------------------------------------------

#### Figure 1.3: QQ Plot of Residuals - Initial Model

```{r figure1.3, echo = FALSE}
# Notice that each code chunk is separated by a blank line, ***, and another blank line

plot(lm_initial, which = 2)
```

------------------------------------------------------------------------

#### Figure 1.4: Histogram - Initial Model (with a log transformation)

```{r figure1.4, echo = FALSE}
# Feel free to use grid.arrange (from the gridExtra package) to plot before and after plots next to each other (but you don't have to!) Here's an example if you're interested
p1 <- ggplot(data = cancer_mortality_rates, aes(x = cmRate)) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of cmRate",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")

p2 <- ggplot(data = cancer_mortality_rates, aes(x = log(cmRate))) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of cmRate",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1, ncol = 2)


p1 <- ggplot(data = cancer_mortality_rates, aes(x = medianAge)) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of medianAge",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")

p2 <- ggplot(data = cancer_mortality_rates, aes(x = log(medianAge))) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of medianAge",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1, ncol = 2)

p1 <- ggplot(data = cancer_mortality_rates, aes(x = pctBach)) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of pctBach",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")

p2 <- ggplot(data = cancer_mortality_rates, aes(x = log(pctBach))) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of pctBach",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1, ncol = 2)


p1 <- ggplot(data = cancer_mortality_rates, aes(x = pctPoverty)) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of pctPoverty",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")

p2 <- ggplot(data = cancer_mortality_rates, aes(x = log(pctPoverty))) + 
  geom_histogram(bins = 30, fill = "grey80", color = "black") + 
  labs(title = "Histogram of pctPoverty",
       subtitle = "by Stats 401 Group34",
       x = "number of cancer mortalities per 100,000 people",
       y = "Frequency")
grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

------------------------------------------------------------------------

#### Figure 1.5: Scatterplot Matrix of Data - Initial Model

```{r figure1.5, echo = FALSE}
plot(~ cmRate + medianAge + pctPoverty + pctBach + rmRace, data = cancer_mortality_rates)
plot(~ cmRate + medianAge + log(pctPoverty) + log(pctBach) + rmRace, data = cancer_mortality_rates)
```

------------------------------------------------------------------------

#### Figure 1.6: Scatterplot - Initial Model

```{r figure1.6, echo = FALSE}
p1 <- ggplot(data = cancer_mortality_rates, aes(x = pctPoverty, y = cmRate)) + 
  
  geom_point() + 
  
  labs(title = "Scatterplot of pctPoverty vs cmRate",
       subtitle = "by Stats 34 group",
       x = "percent of country living in poverty",
       y = "number of cancer moralities per 100,000 people")

p2 <- ggplot(data = cancer_mortality_rates, aes(x = log(pctPoverty), y = cmRate)) + 
  
  geom_point() + 
  
  labs(title = "Scatterplot of log(pctPoverty) vs cmRate",
       subtitle = "by Stats 34 group",
       x = "log(percent of country living in poverty)",
       y = "number of cancer moralities per 100,000 people")
grid.arrange(p1, p2, nrow = 1, ncol = 2)


p1 <- ggplot(data = cancer_mortality_rates, aes(x = pctBach, y = cmRate)) + 
  
  geom_point() + 
  
  labs(title = "Scatterplot of pctBach vs cmRate",
       subtitle = "by Stats 34 group",
       x = "percent of county residents (ages 25+) with a bachelor’s degree",
       y = "number of cancer moralities per 100,000 people")

p2 <- ggplot(data = cancer_mortality_rates, aes(x = log(pctBach), y = cmRate)) + 
  
  geom_point() + 
  
  labs(title = "Scatterplot of log(pctBach) vs cmRate",
       subtitle = "by Stats 34 group",
       x = "log(percent of county residents (ages 25+) with a bachelor’s degree)",
       y = "number of cancer moralities per 100,000 people")
grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

------------------------------------------------------------------------

#### PART 2 : Figure 2.1 : Summary Ouput - Updated Model 1 (with log transformation)

```{r figure2.1, echo = FALSE}
lm_initial1 <- lm(cmRate ~ medianAge + log(pctPoverty) + log(pctBach) + rmRace , data = cancer_mortality_rates)
summary(lm_initial1)
```

------------------------------------------------------------------------

#### Figure 2.2: Residuals plot - Updated Model 1

```{r figure2.2, echo = FALSE}
plot(lm_initial1, which = 1)
```

------------------------------------------------------------------------

#### Figure 2.3: QQ Plot of Residuals - Updated Model1

```{r figure2.3, echo = FALSE}

plot(lm_initial1, which = 2)
```

------------------------------------------------------------------------

#### PART 3 : Figure 3.1: Scatterplot with Color - Initial Model

```{r figure3.1, echo = FALSE}
ggplot(data = cancer_mortality_rates, aes(x = pctBach, y = cmRate)) + 
  
  geom_point(aes(color = rmRace, shape = rmRace)) + 
  
  labs(title = "Scatterplot of pctBach vs cmRate",
       subtitle = "group 34",
       x = "percent of county residents (ages 25+) with a bachelor's degree ",
       y = "number of cancer moralities per 100,000 people") + 
  
  scale_color_manual(values = c("darkblue", "orange")) +
  scale_shape_manual(values = c(15, 16)) + 
  
  theme_bw() 

ggplot(data = cancer_mortality_rates, aes(x = pctPoverty, y = cmRate)) + 
  
  geom_point(aes(color = rmRace, shape = rmRace)) + 
  
  labs(title = "Scatterplot of pctPoverty vs cmRate",
       subtitle = "group 34",
       x = "percent of country living in poverty ",
       y = "number of cancer moralities per 100,000 people") + 
  
  scale_color_manual(values = c("darkblue", "orange")) +
  scale_shape_manual(values = c(15, 16)) + 
  
  theme_bw() 

ggplot(data = cancer_mortality_rates, aes(x = medianAge, y = cmRate)) + 
  
  geom_point(aes(color = rmRace, shape = rmRace)) + 
  
  labs(title = "Scatterplot of medianAge vs cmRate",
       subtitle = "group 34",
       x = "median age of country residents",
       y = "number of cancer moralities per 100,000 people") + 
  
  scale_color_manual(values = c("darkblue", "orange")) +
  scale_shape_manual(values = c(15, 16)) + 
  
  theme_bw() 
```

------------------------------------------------------------------------

#### PART 4 : Figure 4.1 Updated Model 2 (remove rmRace and add region)

```{r figure4.1, echo = FALSE}

lm_initial2 <- lm(cmRate ~ medianAge + pctPoverty + pctBach + region, data = cancer_mortality_rates)
summary(lm_initial2)
```

------------------------------------------------------------------------

#### Figure 4.2: Residuals Plot - Updated Model2

```{r figure4.2, echo = FALSE}
plot(lm_initial2, which = 1)
```

------------------------------------------------------------------------

#### Figure 4.3: QQ Plot of Residuals - Updated Model2

```{r figure4.3, echo = FALSE}

plot(lm_initial2, which = 2)
```

------------------------------------------------------------------------

#### Figure 4.4: Scatterplot with Color - Updated Model2

```{r figure4.4, echo = FALSE}
# pctBach vs cmRate
ggplot(data = cancer_mortality_rates, aes(x = pctBach, y = cmRate)) + 
  geom_point(aes(color = region, shape = region)) + 
  geom_smooth(aes(color = region), method = "lm", se = FALSE) +  
  labs(title = "Scatterplot of pctBach vs cmRate",
       subtitle = "group 34",
       x = "percent of county residents (ages 25+) with a bachelor's degree ",
       y = "number of cancer moralities per 100,000 people") + 
  scale_color_manual(values = c("darkblue", "orange", "yellow", "green", "purple")) +
  scale_shape_manual(values = c(15, 16, 17, 18, 19)) + 
  theme_bw()

# pctPoverty vs cmRate
ggplot(data = cancer_mortality_rates, aes(x = pctPoverty, y = cmRate)) + 
  geom_point(aes(color = region, shape = region)) + 
  geom_smooth(aes(color = region), method = "lm", se = FALSE) +  
  labs(title = "Scatterplot of pctPoverty vs cmRate",
       subtitle = "group 34",
       x = "percent of country living in poverty",
       y = "number of cancer moralities per 100,000 people") + 
  scale_color_manual(values = c("darkblue", "orange", "yellow", "green", "purple")) +
  scale_shape_manual(values = c(15, 16, 17, 18, 19)) + 
  theme_bw()

# medianAge vs cmRate
ggplot(data = cancer_mortality_rates, aes(x = medianAge, y = cmRate)) + 
  geom_point(aes(color = region, shape = region)) + 
  geom_smooth(aes(color = region), method = "lm", se = FALSE) +  
  labs(title = "Scatterplot of medianAge vs cmRate",
       subtitle = "group 34",
       x = "median age of country residents",
       y = "number of cancer moralities per 100,000 people") + 
  scale_color_manual(values = c("darkblue", "orange", "yellow", "green", "purple")) +
  scale_shape_manual(values = c(15, 16, 17, 18, 19)) + 
  theme_bw()

```

------------------------------------------------------------------------

#### Figure 4.5: Interaction - Updated Model2

```{r figure4.5, echo = FALSE}
lm_initial2 <- lm(cmRate ~ medianAge + pctPoverty + pctBach + region, data = cancer_mortality_rates)
summary(lm_initial2)

lm_interaction1 <- lm(cmRate ~ medianAge + pctPoverty + pctBach + region + region * medianAge, data = cancer_mortality_rates)
summary(lm_interaction1)

lm_interaction2 <- lm(cmRate ~ medianAge + pctPoverty + pctBach + region + region * pctPoverty, data = cancer_mortality_rates)
summary(lm_interaction2)

lm_interaction3 <- lm(cmRate ~ medianAge + pctPoverty + pctBach + region + region * pctBach, data = cancer_mortality_rates)
summary(lm_interaction3)
```

------------------------------------------------------------------------

#### PART 5 (Final Model) : Figure 5.1

```{r figure5.1}

lm_final <- lm(cmRate ~ medianAge + log(pctPoverty) + log(pctBach) + region + medianAge * region + log(pctPoverty) * region, data = cancer_mortality_rates)
summary(lm_final)

```

------------------------------------------------------------------------

#### Figure 5.2: Residuals Plot - final Model

```{r figure5.2, echo = FALSE}
plot(lm_final, which = 1)
```

------------------------------------------------------------------------

#### Figure 5.3: QQ Plot of Residuals - Final Model

```{r figure5.3, echo = FALSE}
# Notice that each code chunk is separated by a blank line, ***, and another blank line

plot(lm_final, which = 2)

```

------------------------------------------------------------------------

#### Figure 5.4: VIF - Final Model

```{r figure5.4, echo = FALSE}

vif(lm(cmRate ~ medianAge + log(pctPoverty) + log(pctBach), data = cancer_mortality_rates))

```

------------------------------------------------------------------------

#### Final Model - CMF

$E(Y_{cmRate}|X) = \beta_0 + \beta_1 X_{medianAge} + \beta_2 log(X_{pctPoverty}) + \beta_3 log(X_{pctBach}) + \beta_4 I_{Northeast} + \beta_5 I_{Southeast} + \beta_6 I_{Southwest} + \beta_7 I_{West} + \beta_8 I_{Northeast} * X_{medianAge} + \beta_9 I_{Southeast} * X_{medianAge} + \beta_{10} I_{Southwest} * X_{medianAge} + \beta_{11} I_{West} * X_{medianAge} + \beta_{12} I_{Northeast} * log(X_{pctPoverty}) + \beta_{13} I_{Southeast} * log(X_{pctPoverty}) + \beta_{14} I_{Southwest} * log(X_{pctPoverty}) + \beta_{15} I_{West} * log(X_{pctPoverty})$

